EuPPollNet: A European database of plant-pollinator networks

### 2024-05-23 Manuscript/EuPPollNet_Ms.qmd
edits

p1 = ggplot(c2, 
aes(fill=Pollinator_order, y=Proportion, x=Habitat)) + 
geom_bar(position = position_stack(reverse = TRUE), stat="identity")+
theme_bw() +
theme(legend.position = "bottom",
      plot.margin=unit(c(0,0,0,0), "mm"),
      legend.text = element_text(size=6),
      legend.key.size = unit(2, "mm"),
      legend.spacing.y = unit(0.1, "cm"),
       legend.box.spacing = unit(0, "pt"),
      axis.title = element_text(face = "bold", size = 7),
      axis.text = element_text(size=6),
      panel.border = element_rect(size=1)) +
xlab(NULL) +
scale_y_continuous(breaks = c(0.25, 0.5,0.75), labels = c("0.25", "0.5","0.75"))+
theme(strip.text.x = element_text(size = 10, face = "bold"),
    strip.background = element_rect(
color=NA, fill=NA, size=0.5, linetype="solid")) +
scale_fill_manual(values = viridis(4),
                  name = NULL) +
coord_flip(expand = FALSE) +
ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
guides(fill=guide_legend(nrow=2,byrow=TRUE))

@@

Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and as expected, followed an negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**). Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and as expected for this metric, were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S3**). We found that only `r round(over_z_percent, 2)`% of networks were statistically more nested than expected by chance, with `r round(no_diff_z_percent, 2)`% showing no statistical difference, and none being less nested than null expectations (**Figure 6b**). Latitude explained little of the observed variability of residual connectance and nestedness across networks (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**). Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness that networks located in lower latitudes.

@@

Although Europe contains a much larger number of flowering plants than pollinator species (~4 to 1 ratio), the observed number of pollinator species almost doubled the one of plants in the database. This could be explained by the fact that all networks are phytocentric, resulting in sampling bias towards pollinator species [@jordano2016; @vizentin2018]. While animal-centred sampling is likely to increase the plant-pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants. As expected [@jordano1987], we found that network connectance decreased exponentially with the number of species. The less connected networks were the Alpine ones as they were the ones with higher number of species.

#Task: Calcular residual connectance.

with more species had lower connectance .

Finish this paragraph with an idea of how sampling can bias our view and that is important to be aware of that when understanding plant-pollinator patterns

Discuss that other type of informative networks could be useful an increase connectance?

visitation networks

@@

Nestedness
Pattern across latitude

Missing links?

>>>>

p1 = ggplot(c2, 
aes(fill=Pollinator_order, y=Proportion, x=Habitat)) + 
geom_bar(position = position_stack(reverse = TRUE), stat="identity")+
theme_bw() +
theme(legend.position = "bottom",
      plot.margin=unit(c(0,0,0,0), "mm"),
      legend.text = element_text(size=6),
      legend.key.size = unit(2, "mm"),
      legend.spacing.y = unit(0.1, "cm"),
       legend.box.spacing = unit(0, "pt"),
      axis.title = element_text(face = "bold", size = 7),
      axis.text = element_text(size=6),
      panel.border = element_rect(size=1)) +
xlab(NULL) +
ylab("Proportion of species") +
scale_y_continuous(breaks = c(0.25, 0.5,0.75), labels = c("0.25", "0.5","0.75"))+
theme(strip.text.x = element_text(size = 10, face = "bold"),
    strip.background = element_rect(
color=NA, fill=NA, size=0.5, linetype="solid")) +
scale_fill_manual(values = viridis(4),
                  name = NULL) +
coord_flip(expand = FALSE) +
ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
guides(fill=guide_legend(nrow=2,byrow=TRUE))

@@

Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and as expected, followed an negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**). Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and as expected for this metric, were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S3**). We found that only `r round(over_z_percent, 2)`% of networks were statistically more nested than expected by chance, with `r round(no_diff_z_percent, 2)`% showing no statistical difference, and none being less nested than null expectations (**Figure 6b**). Latitude explained little of the observed variability of residual connectance and nestedness across networks (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6a**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6b**). Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness that networks located in lower latitudes. Note that residual connectance and normalised nestedness have a moderate negative correlation (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`).

@@

Although Europe contains a much larger number of flowering plants than pollinator species (~4 to 1 ratio), the observed number of pollinator species almost doubled the one of plants in the database. This could be explained by the fact that all networks are phytocentric, resulting in sampling bias towards pollinator species [@jordano2016; @vizentin2018]. While animal-centred sampling is likely to increase the plant:pollinator species ratio [e.g., @encinas2023], the spatial scale and environmental context of the sampled communities will also influence their observed diversity, especially given the ability to move of pollinators and the sessile nature of plants. As expected [@jordano1987], we found that network connectance decreased exponentially with the number of species. Overall, latitude did not explain major changes in residual connectance or normalised nestedness of the plant-pollinator networks. Consistent with @olesen2002, residual connectance was lower at higher latitudes, while normalised nestedness, which is negatively correlated with residual connectance, increased towards higher latitudes. Given their correlation, the potential ecological explanations that can help understanding these structural patterns can be similar. For instance, networks at lower latitudes in Europe are exposed to higher temperatures and are bee-dominated, which can result in higher visitation rates [@arroyo1985; @classen2015; @herrera2019]; and, the level of pollinator generalization seems to be higher at lower latitudes [@schleuning2012]. Therefore, these different factors can increase the possible number of connections that can be established between plants and pollinators for a given network size, resulting in more connected networks but less nested ones in comparison to networks at higher latitudes. In addition, while one third of networks were more nested than expected by chance, two thirds of them did not show statistical difference with null expectations. This goes against the common view that all plant-pollinator networks are nested [@bascompte2003], and supports the idea that despite a strong tendency towards it, being nested is not the rule in this type of mutualism.

Despite this database covers a wide range of habitats across Europe, still contains major geographical gaps such as the East of Europe or the Mediterranean region.

Maybe some refs about the above phrase?

Add that

--> Methodology, impact the structure of the network

Finally, these networks have different sampling effort and employ different metodologies which can impact the resulting structure of interactions [REF].

Finish this paragraph with an idea of how sampling can bias our view and that is important to be aware of that when understanding plant-pollinator patterns
Discuss that other type of informative networks could be useful a increase connectance?

@@
